Liveness-Based Garbage Collection for Lazy Languages

### 2015-06-05 paper/paper.tex
Adding terminology to implementation

In this  section we describe in  detail the process  of liveness based
garbage collection.  During  the execution of a program  if it runs of
out  of memory,  execution is  stopped  and the  garbage collector  is
invoked. At this point all the references on the stack are checked for
liveness and only live references are copied and the program continues
execution.  The main  complexity in our algorithm is  the way closures
are  handled. The  following sub-sections  show two  different  ways of
handling closures along with their pros and cons.
\subsection{A simple LGC Algorithm}
\cred{Introduce terminology}
% Abstract description of the GC algorithm
\begin{verbatim}
proceudre lgc()
 FOREACH reference in root set
 DO
   IF reference is live
     copy(cell)
   ENDIF
 DONE

@@

\subsection{Modified GC Algorithm}
Since the first algorithm requires a lot of bookkeeping and is not
precise, we have a second version of the algorithm 
which does not use liveness for copying closures. 
\begin{verbatim}
procedure copy (cell)
  IF{celltype is cons}
    IF cell is live &&
         cell is not copied using reachability
       new_cell <-  dup_heapcell(cell) 
       new_car        <-  copy(cell.car)
       new_cell.car   <-  new_car
       new_cdr        <-  copy(cell.cdr)
       new_cell.cdr   <-  new_cdr
    ELSE
       cell <- NULL
    ENDIF
  ELSE
    IF {celltype is closure && cell is not copied}
       new_cell      <-  dup_heapcell(cell)
       new_cell.arg1 <-  copy(cell.arg1)
       new_cell.arg2 <-  copy(cell.arg2)
    ENDIF
  ENDIF
 copy_references_on_print_stack()
\end{verbatim}

>>>>

In this  section we describe in  detail the process  of liveness based
garbage collection.  During  the execution of a program  if it runs of
out  of memory,  execution is  stopped  and the  garbage collector  is
invoked. At this point all the references on the stack are checked for
liveness and only live references are copied and the program continues
execution.  The main  complexity in our algorithm is  the way closures
are  handled. The  following sub-sections  show two  different  ways of
handling closures along with their pros and cons.
\subsection{A simple LGC Algorithm}
\cred{Introduce terminology}
We introduce certain terminology that we use in our discussion of the 
garbage collection algorithms. A reference is a field which holds the 
address of a heap location. Root set refers to the set of all references 
that are on the program stack during garbage collection.
A heap cell can either contain an atomic value, a \CONS\ cell or a closure. A closure 
has fields to identify the closure type and 2 arguments. In case of function calls 
with more than 2 arguments we use currying (one of the argument fields will point to another
 heap location containing the next argument and so on). By reachability we mean all heap 
locations that are reachable traversing the argument fields of a given reference.
% Abstract description of the GC algorithm
\begin{verbatim}
proceudre lgc()
 FOREACH reference in root set
 DO
   IF reference is live
     copy(cell)
   ENDIF
 DONE

@@

\subsection{Modified GC Algorithm}
Since the first algorithm requires a lot of bookkeeping and is not
precise, we have a second version of the algorithm 
which does not use liveness for copying closures. 
\begin{verbatim}
procedure copy (cell)
  IF{celltype is cons}
    IF cell is live &&
         cell is not copied using reachability
       new_cell <-  dup_heapcell(cell) 
       new_car        <-  copy(cell.car)
       new_cell.car   <-  new_car
       new_cdr        <-  copy(cell.cdr)
       new_cell.cdr   <-  new_cdr
    ELSE
       cell <- NULL
    ENDIF
  ELSE
    IF {celltype is closure && cell is not copied}
       new_cell  <- copy_and_mark_reachable_cells(cell)
    ENDIF
  ENDIF
 copy_references_on_print_stack()
\end{verbatim}
